Buildtime Trend

### 2016-01-26 buildtimetrend/btt_fosdem2016.tex
start working on Fosdem presentation

\title[Buildtime Trend : What, why and how]{Buildtime Trend : What's trending in your build process}
\subtitle{What, why and how}

@@

\date[NewLine 0x05 5Apr2015]{NewLine 0x05 - Ghent\\
April 5th, 2015}

@@

      \item What is Buildtime Trend
      \item Why and how

@@

      \item Lessons learned

@@

    \pause
    \textbf{Solution:} create a script to generate timestamps
    \frametitle{Next step : calculate duration and generate chart}
    Requirements :
      \item parse the generated timestamps
      \item calculate duration of the stages
      \item store the timing data of every build
      \item generate the chart
    \pause
    One solution : Bash + CSV + gnuplot\\
    \pause
    Another solution : Python + XML + matplotlib
    \frametitle{Problem}
    I didn't use Python before.\\
    \pause
    A good opportunity to learn Python!
    \frametitle{Learning Python}
      \item start with a tutorial
      \item read documentation
      \item ask Google and Stack Overflow
      \item read code from other projects
      \item get advice from friends and colleagues
      \item go to talks and conferences
    \frametitle{Other helpful things}
      \item check coding style
          \item commandline : pep8, pylint
          \item online tools : Landscape.io, Scrutinizer, Codacy
      \item unit testing and test driven development (TDD)
      \item code coverage
      \item automate this with Continuous Integration (CI) : Travis CI
      \item version control : Git, GitHub, ...

@@

    \frametitle{Limitations}

@@

      \item scalability of XML
      \item querying data is limited
      \item developing new charts is not efficient

@@

    \frametitle{Integration with Travis CI}
    Call Buildtime Trend scripts in .travis.yml file
      \item generate timestamps
      \item install dependencies
      \item analyze timestamps and generate chart
      \item store timing data (XML) and chart in gh-pages
    \frametitle{Integration with Travis CI}
    This is not ideal!
      \item scripts become part of the build process
      \item build duration is influenced
      \item complicated to set up
    \frametitle{Integration with Travis CI}
    Possible solutions
      \item create a service that analyses the logfile after the build is finished
      \item store the data in a database outside Travis
      \item host the generated chart(s) outside Travis
    But how?

@@

      \item gather data
      \item store it
      \item analyse
      \item visualise
      \item support for several languages, including Python
    Great : this solves the data storage and chart generation problem!

@@

    \frametitle{First release - features}
      \item scripts tightly coupled with Travis CI build
      \item storing data in XML and chart generation with matplotlib (native mode)
      \item storing data and chart generation with Keen.io
      \item chart dashboard deployed to gh-pages
    Yeah, it works!\\
    \pause
    But there is still room for improvement.
    \frametitle{Travis CI timing data}
    Good news!\\
    Travis CI announces adding timing data to their logs!\\
    \frametitle{Create a service}
      \item parse the timing data
      \item store data in Keen.io
    \frametitle{Rearrange project structure}
    Create organisation on GitHub to host the different repositories :\\
    \includegraphics[scale=.5]{buildtimetrend_diagram.png}
    \frametitle{Challenges}
      \item let the subprojects work together
      \item Open Source license
      \item triggering the service at the end of a build
      \item business model
    \frametitle{Combine subprojects}
    \item Client and service use dashboard : git submodule
    \item Client and service use library :
        \item create Python package \textit{buildtimetrend}
        \item available on \textit{pypi}
        \item install as dependency
    \frametitle{Open Source license}
    Transition from GNU General Public License (GPL) v3 to Affero GPLv3

@@

          \item badge

@@

      \item deploy to Heroku with Travis CI build

@@

    \frametitle{Deploy to Heroku during Travis CI build}
    In \textit{.travis.yml} :
      \small{deploy:\\
      \ \ provider: heroku\\
      \ \ api\_key:\\
      \ \ \ \ secure: <secure key>\\
      \ \ app:\\
      \ \ \ \ deploy-prod: buildtimetrend}
    Branch \textit{deploy-prod} is deployed to app \href{https://buildtimetrend.herokuapp.com/}{https://buildtimetrend.herokuapp.com/}
    \frametitle{Travis CI triggers service}

@@

    \frametitle{Travis CI triggers service}
      \item a JSON payload with build data is sent
      \item service identifies repo and build ID
      \item build job data and logfiles are retrieved from Travis CI
      \item build job logfiles are parsed looking for timing tags
      \item build stages duration is calculated
      \item build job data is sent to Keen.io

@@

      \item service : seperate worker for build job processing

@@

    \frametitle{Lessons learned}
      \item start simple
      \item (re)use as much existing tools as possible
      \item start with what you know, learn new skills as you progress
      \item try, even if it is all scary at first

@@

      \item Josh (\href{https://twitter.com/dzello}{@dzello}) and the other nice people of Keen.io, for their invaluable support!
      \item \href{https://twitter.com/thomasbonte}{@thomasbonte} for introducing me to Lean Startup principles
      \item Alex for telling me about CherryPy

@@

      \item my friends and family

>>>>

\title[Buildtime Trend]{Buildtime Trend}
\subtitle{What's trending in your build process}

@@

\date[FOSDEM 2016]{FOSDEM 2016 - Brussels\\
January 31st, 2016}

@@

      \item What is Buildtime Trend?
      \item How does it work?
      \item How to start using it

@@

@@

@@

    \frametitle{How it works}
    Service :

@@

      \item parse the (timing) data
      \item store data
      \item visualise using a dashboard with trends

@@

@@

@@

@@

          \item badges

@@

@@

    \frametitle{How to use it}

@@

@@

@@

@@

      \item The nice people of Keen.io, for their invaluable support!

@@
